\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage[margin=1.5cm]{geometry}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{xparse}
\usepackage{enumitem, inconsolata}
\parindent 0px
\newcommand{\lb}{\\$\left|\rightarrow\right.$}
\newcommand{\enter}{\\\textcolor{white}{1}}

\ExplSyntaxOn
\NewDocumentCommand{\bo}{m}
 {
   \bold_commas:n { #1 }
 }

\cs_new:Npn \bold_commas:n #1
 {
   \seq_set_split:Nnn \l_tmpa_seq { , } { #1 }
   \seq_map_indexed_function:NN \l_tmpa_seq \__bold_commas_aux:nn
 }

\cs_new:Npn \__bold_commas_aux:nn #1 #2
 {
   \textbf{#2}
   \int_compare:nNnTF { #1 } < { \seq_count:N \l_tmpa_seq }
     { , }
     { }
 }

\ExplSyntaxOff

\title{Computer Network}
\author{Me lol}
\date{\today}

\begin{document}
\maketitle
\vspace{13cm}
\begin{large}\textbf{Notes}\end{large}
\begin{itemize}
\item PYQs of BEX's CT657, BEI's CT613 and BCT's CT702 are combined.
\item BEX's and BCT's are kept with normal font.\item BEI's are kept with \texttt{this styling to differentiate}.
\end{itemize}
\pagebreak
\tableofcontents
\pagebreak

\section{Introduction to Computer Network}
	\begin{center}(5 Hours/8 Marks)\end{center}
	\subsection{Computer Network \& Uses}
		\begin{enumerate}[noitemsep, topsep = 0pt]
			\item What are the applications of computer networks?\hfill[2] (76 Ba)
			\item What is computer network?\hfill[1] (76 Ba) [2] (71 Shr)
			\item What are the advantages of computer network?\hfill[3] (72 Ma)
			\item Explain five instances of how networks are a part of your life today. \hfill [5] (\bo{72 Ch})
		\end{enumerate}

	\subsection{Networking model: client/server, P2P, active network}
		\begin{enumerate}[noitemsep, topsep = 0pt]
			\item Discuss Client-Server model and Peer-to-Peer model.\hfill[4] (\texttt{81 Ba})
			\item How does client-server model work?\hfill[3] (73 Ma)
			\item What are the features of Client/Server Architecture?\hfill[4] (70 Ch, 76 Ash)
			\item Draw the architecture for Client/Server network model.\hfill[2] (\bo{75 Ch})
			\item Explain briefly the architecture for peer-to-peer network model with example.\hfill[3] (\bo{77 Ch})
			\lb How does P2P works? Explain.\hfill[6] (\bo{\texttt{80 Bh}})
			\lb Explain in details about P2P network model with supportive examples.\hfill[6] (\bo{75 Ch})
			\item Distinguish between Client-Server network and Peer-Peer network.\hfill[3] (\bo{74 Ch}) [5] (73 Ma)
		\end{enumerate}

	\subsection{Network Software, Protocols and Standards}
		\begin{enumerate}[noitemsep, topsep = 0pt]
			\item Define protocol.\hfill[1] (\bo{76 Ch, 72 Ash}) [4] (bo{79 Ch}) 

			\item Explain about connection oriented and connection less service.\hfill[3] (\bo{71 Bh})
			\item Mention service primitives for implementing connection oriented service.\hfill[2] (75 Ash)
			\item What do you mean by network architecture?\hfill[2] (\bo{71 Ch})
			\item Why do you need layering?\hfill[2] (\bo{\texttt{80 Bh}})
			\item Why do we need layered protocol architecture?\hfill[2] (\bo{72 Ash})
			\item What are the reasons for using layered network architecture?\hfill[2] (\bo{76 Ch}) [3] (\bo{73 Ch}, 75 Ba)
			\item Why layering is important?\hfill[2] (\bo{\texttt{79 Bh}}, 75 Ash) [4] (\bo{73 Bh})
			\item What are the layer design issues?\hfill[3] (\bo{71 Bh})
			\item Explain design issues for layers in detail.\hfill[4] (75 Ash)
			\item Explain about the design issues of Computer Network software.\hfill[5] (\bo{74 Bh})
		\end{enumerate}

	\subsection{OSI model and TCP/IP model}
		\begin{enumerate}[noitemsep, topsep = 0pt]
			\item Explain Open System Interconnection (OSI) model.\hfill[5] (\bo{74 Ch}, 71 Ma)
			\lb Explain OSI reference model with suitable diagram. \hfill [6] (\texttt{80 Ba})
			\lb Explain the different layers of OSI reference Model with appropriate figure.\hfill[5] (\bo{76 Bh})
			\item What is the significance of OSI layer?\hfill[2] (74 Ash)
			\lb What are the reasons for using layered protocol?\hfill[3] (75 Bh)
			\item Explain in which level of OSI layer following tasks are done.\hfill[3] (\bo{72 Ch})\\
			(i) Error detection and correction \hspace{20mm} (ii) Encryption and Decryption of data\\
			(iii) Logical identification of computer \hspace{11mm} (iv) Point-to-point connection of socket\\
			(v) Dialogue control \hspace{45mm}(vi) Physical identification of computer
			\item Explain each layer of TCP/IP protocols architecture in detail.\hfill[5] (\bo{72 Ash})

			\item List out the functions of physical lapyer in TCP/IP reference model. \hfill [2] (72 Ka)

			\item Distinguish between physical channel and physical layer.\hfill[3] (\bo{74 Bh})
		\end{enumerate}

		\subsection{Comparison of OSI and TCP/IP model}
			\begin{enumerate}[noitemsep, topsep = 0pt]
				\item Differentiate between TCP/IP and OSI Model.
				\enter\hfill[3] \begin{footnotesize}(\bo{76 Bh, 71 Ch}, 71 Ma)\end{footnotesize} [4] \begin{footnotesize}(81 Ba)\end{footnotesize} [5] \begin{footnotesize}(\bo{77 Ch, 76 Ch, 72 Ch}, 76 Ba, 75 Ba, 73 Shr, 72 Ma)\end{footnotesize} [6] \begin{footnotesize}(71 Shr)\end{footnotesize} 
				\lb Explain OSI model and compare OSI with TCP/IP reference model.\hfill[6] (\bo{\texttt{79 Bh}})
				\item How do you differentiate OSI's network and transport layer with TCP/IP's Network and transport layer?\hfill[4] (\bo{79 Ch})
			\end{enumerate}

		\subsection{Data Encapsulation}
		\begin{enumerate}[noitemsep, topsep = 0pt]
			\item What is data encapsulation? \hfill [2] (\texttt{80 Ba})
			
			\item How the process of data encapsulation occurs in transmission mode described by seven layers of OSI model.\hfill[5] (75 Bh)

			\item What are headers and trailers and how do they get added and removed?
		\enter\hfill[4] (76 Ash, 70 Ch) [5] (\bo{73 Ch}) 
		\end{enumerate}

		\subsection{Example network: The Internet, X.25, Frame Relay, Ethernet, VoIP, NGN and MPLS, xDSL}
		\begin{enumerate}[noitemsep, topsep = 0pt]
		\item Explain X.25 Network with its key feature.\hfill[3] (\bo{71 Ch})
		\item Define Frame Relay in detail.\hfill[3] (73 Shr)
		\item What is Internet work?\hfill[2] (\bo{71 Bh})
		\item Explain the function of following of following devices in brief:\hfill[6] (\bo{73 Bh})\\
		(i) Hub (ii) Bridge (iii) Router
		\item You are assigned to design a network infrastructure for a 3-start hotel. Recommend a network solution with hardwares and softwares in current trend that can be used in the hotel. Make necessary assumptions and justify your recommendation with logical arguments where possible.\hfill[8] (72 Ka)
		\end{enumerate}

	\pagebreak

\section{Physical Layer}
	\begin{center}(5 Hours/8 Marks)\end{center}
	\subsection{Network monitoring: delay, latency, throughput}
		\begin{enumerate}[noitemsep, topsep=0pt]	
			\item Define:
			\lb Delay \hfill [1] (\bo{\texttt{80 Bh}}, \texttt{80 Ba})
			\lb Bandwidth \hfill [1] (\bo{\texttt{80 Bh}, 76 Bh})
			\lb MAC address \hfill [1] \bo{\texttt{80 Bh}}
			\lb Throughput \hfill [1] (\bo{76 Bh}, \texttt{80 Ba}, 76 Ba)
			\lb Latency \hfill [1] (76 Ba)
			
			\item What are the causes of packet delay? \hfill [2] (\bo{76 Bh})
		\end{enumerate}

	\subsection{Transmission media: Twisted pair, Coaxial, Fiber optic, Line-of-site, Satellite}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item What is transmission medium?  \hfill [1] (\bo{76 Ch, 71 Bh}) [2] (71 Shr) [3] (\bo{74 Ch})
			
			\item Write down the transmission medium used for networking. \hfill [4] (\bo{72 Ash})
			
			\item Explain different types of transmission media. \hfill [6] (72 Ka)
			\lb Explain about any three transmission media in detail. \hfill [6] (71 Shr)
			
			\item What are the factors to be considered while selecting transmission media? \hfill [2] (\bo{\texttt{79 Bh}}, \texttt{81 Ba})

			\item Explain twisted pair cable with its practical applications. \hfill [4] (\bo{77 Ch})	
			
			\item Explain optical fiber cable in detail with its advantages and disadvantages. \hfill [6] (\texttt{81 Ba})
			
			\item Explain different transmission medium with their merits and demerits. \hfill [6] (\bo{76 Ch})
			
			\item Explain the characteristics of twisted pair, coaxial and optical fiber cable. \hfill [6] (76 Ba)
			
			\item Explain about any two guided transmission media in detail. \hfill [6] (74 Ash)	
			
			\item List guided and unguided media used in computer network. \hfill [2] (73 Ma)
			
			\item Explain Ethernet cable standards. \hfill [6] (73 Ma)
			
			\item Compare different types of guided transmission media with appropriate figures. \hfill [6] (\texttt{80 Ba})
			
			\item Differentiate between wired and wireless media with their benefits and drawbacks. \hfill [3] (75 Ba)
			
			\item Compare among Twisted Pair, Coaxial cable and Fiber optic. \hfill [5] (\bo{74 Ch})
			
			\item Why, now a day all communication media like twisted pair, co-axial pair even wireless media are replaced by optical fibre? Justify your answer with necessary diagram, working principle and transmission mechanism. \hfill [7] (\bo{71 Bh})
			
			\item Explain various cabling techniques used in IEEE 802.3 standard. \hfill [4] (71 Ma)
		\end{enumerate}

	\subsection{Multiplexing, Circuit switching, Packet switching, VC Switching, Telecommunication switching system (Networking of Telephone exchanges)}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item What is switching? \hfill [1] (\bo{75 Ch}, 74 Ash) [2] (\bo{74 Bh}, 73 Shr)[3] (\bo{75 Bh})
			
			\item What are the various switching techniques? \hfill [2] (\bo{75 Ch})
			\lb Explain about various switchings with practical implementation example. \hfill [6] (\bo{70 Ch})
			
			\item What do you mean by data switching? \hfill [2] (\bo{70 Ch})
			
			\item Define multiplexing. \hfill [1] (74 Ash)
			
			\item Explain different types of multiplexing used in communication system. \hfill [4] (72 Ma)
			
			\item Compare switching with multiplexing. \hfill [2] (73 Shr)

			\item Elaborate packet switching with a proper diagram. \hfill [5] (\bo{75 Ch})	
			
			\item (Assumed) Discuss how data or packets goes through switch to switch in Frame Relay Virtual-circuit network. \hspace{12.6cm} [5] (\bo{\texttt{80 Bh}})
			
			\item Differentiate between datagram switching and virtual circuit switching approach. 
			\enter \hfill [4] (72 Ma) [6] (\bo{\texttt{79 Bh}})
			\lb with suitable diagram. \hfill [6] (\bo{74 Bh})
			
			\item Differentiate between circuit switching and packet switching. 
			\enter\hfill [3] (\bo{75 Bh}, 75 Ash) [4] (\bo{77 Ch, 76 Bh, 72 Ash})
			\lb Compare and which would you prefer and why? \hfill [2+2] (71 Ma)
			
			\item Discuss Packet and Circuit switching concepts with example. \hfill [5] (75 Ba)
		\end{enumerate}

	\subsection{ISDN: Architecture, Interface, and Signaling}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item What is ISDN? \hfill [2] (\bo{71 Ch})
			
			\item Explain about the ISDN architecture in detail with example. \hfill [6] (\bo{71 Ch})
			
			\item Why the telephone companies developed ISDN? \hfill [2] (76 Ash)
			
			\item Explain the working principle of ISDN with its interface and functional group. \hfill [6] (76 Ash)
			
			\item Explain ISDN channels with architecture. \hfill [5] (75 Ash)
			
			\item Explain the E1 Telephone hierarchy system. \hfill [4] (73 Shr)
		\end{enumerate}

	\pagebreak

\section{Data Link Layer}
	\begin{center}(5 Hours/8 Marks)\end{center}
	\subsection{Functions of Data link layer}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item How does data link apply flow control technique in network communication? Explain with example.
			\enter\hfill [6] (\bo{79 Ch})

			\item What are the services provided by data link layer? \hfill [2] (\bo{\texttt{80 Bh, 79 Bh}}) [3] (\bo{77 Ch})

			\item What are the different sub-layers of data link layer? \hfill [2] (76 Ba)

			\item Explain the functions of each sub-layer. \hfill [6] (76 Ba)

			\item State the various design issues for the data link layer? \hfill [3] (75 Ash)
		\end{enumerate}
		
	\subsection{Framing}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item What do you mean by Framing? \hfill [2] (\bo{79 Ch})

			\item Compare Flag byte with byte stuff and bit stuffing in Framing. \hfill [2] (\texttt{81 Ba})

			\item Describe the various framing techniques at data link layer. \hfill [2] (\bo{75 Ch}) [6] (\bo{75 Ch})

			\item Explain any one method of framing with example. \hfill [6] (\bo{75 Ch})
		\end{enumerate}

	\subsection{Error Detection and Corrections}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item How can CRC be used to detect error? \hfill [6] (\bo{\texttt{80 Bh}})
			
			\item Detect the error if any using CRC, if received fram is 0101101101 and generator polynomial is 1001. \hfill [6] (\texttt{81 Ba})

			\item Explain Selective repeat and Go back N ARQ with example. \hfill [6] (\texttt{80 Ba})
		\end{enumerate}

	\subsection{Flow Control}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item (Assumed) Differentiate between datagram switching and virtual circuit switching approach.
			\enter\hfill [6] (\bo{\texttt{79 Bh}})

			\item What is piggybacking? \hfill [3] (75 Ash)
		\end{enumerate}

	\subsection{Examples of Data Link Protocol, HDLC, PPP}
	\subsection{The Medium Access Sub-layer}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item (Assumed) Through we have MAC address, why do we use IP address to represent the host in networks? Explain your answer. \hfill [3] (\bo{72 Ch})
		\end{enumerate}

	\subsection{The channel allocation problem}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item Why do you think that static channel assignment is not efficient? \hfill [2] (\bo{73 Ch}) 
		\end{enumerate}

	\subsection{Multiple Access Protocols}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item What are multiple access protocols? \hfill [2] (\bo{76 Bh, 75 Ch})
		\end{enumerate}

	\subsection{Ethernet}
		\begin{enumerate}
			\item Explain Ethernet frame with function of each field. \hfill [5] (\bo{77 Ch})
		\end{enumerate}

	\subsection{Networks: FDDI, ALOHA, VLAN, CSMA/CD, IEEE 802.3(Ethernet), 802.4(Token Bus), 802.5(Token Ring), and 802.1(Wireless LAN)}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item What is ALOHA system? \hfill [2] (\bo{79 Ch})
			\lb Explain different types of ALOHA. \hfill [4] (75 Ba)
			
			\item How does CSMA-CD works? Explain. \hfill [4] (\bo{76 Ch}) [6] (\bo{79 Ch})
			\lb with figure. \hfill [8] (76 Ash)

			\item Why is CSMA-CD not suitable for wireless medium? Explain. \hfill [2] (\bo{76 Ch})
			
			\item Explain about operation of Carrier Sense Multiple Access with Collision Detection. \hfill [6] (\bo{73 Ch})
			
			\item What is collission? How is it occured? \hfill [1+1] (\bo{76 Ch})

			\item How the possibility of collision is reduced in IEEE 802.3 and IEEE 802.11? Explain. \hfill [6] (\bo{76 Ch})

			\item Differentiate between Token Bus and Token Ring networks. \hfill [4] (75 Ba)
		\end{enumerate}
			
	\subsection{Numericals}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item A frame having size of 1000 bits is transmitted through a channel having capacity of 200 KB/Sec. Calculate the percentage of idleness of the channel assuming the round trip time of the channel to be 20ms. Is the channel efficient? What is your recommendation further?
			\enter\hfill [5+1+4] (\bo{73 Bh})

			\item A bit string 0110111011111111011111110 needs to be transmitted with flag 7E at the data linnk layer. What is the string actually transmitted after bit stuffing? \hfill [6] (\texttt{80 Ba})

			\item A bit string 01111011111101111110 needs to be transmitted at the data link layer. What is the string actually transmitted after bit stuffing? \hfill [2] (75 Ash)
		\end{enumerate}

	\pagebreak

\section{Network Layer}
	\begin{center}(9 Hours/16 Marks)\end{center}
	\subsection{Internet working and devices: Repeaters, Hubs, Bridges, Switches, Router, Gateway}
	\subsection{Addressing: Internet address, classful address}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item What are IPv4 address classes? \hfill [2] (\bo{79 Ch})
		\end{enumerate}

	\subsection{Subnetting}
		\begin{enumerate}[noitemsep, topsep=0pt]
			\item How do you subnetwork a class C network block to following subnets of 41,14,102, and 21 computers respectively? Explain with example. \hfill [8] (\bo{79 Ch})
		\end{enumerate}

	\subsection{Routing: techniques, static vs. dynamic routing , routing table for classful address}
	\subsection{Routing Protocols: RIP, OSPF, BGP, Unicast and multicast routing protocols}
	\subsection{Routing algorithms: shortest path algorithm, flooding, distance vector routing, link state routing; Protocols: ARP, RARP, IP, ICMP}

	\pagebreak
\section{Transport Layer}
	\begin{center}(5 Hours/8 Marks)\end{center}
	\subsection{The transport service: Services provided to the upper layers}
	\subsection{Transport protocols: UDP, TCP}
	\subsection{Port and Socket}
	\subsection{Connection establishment, Connection release}
	\subsection{Flow control and buffering}
	\subsection{Multiplexing and de-multiplexing}
	\subsection{Congestion control algorithm: Token Bucket and Leaky Bucket Transport Layer}

	\pagebreak

\section{Application Layer}
	\begin{center}(5 Hours/8 Marks)\end{center}
	\subsection{Web: HTTP and HTTPS}
	\subsection{File Transfer: FTP, PuTTY, WinSCP}
	\subsection{Electronic Mail: SMTP, POP3, IMAP}
	\subsection{DNS}
	\subsection{P2PApplications}
	\subsection{Socket Programming}
	\subsection{Application server concept proxy caching, Web/Mail/DNS server optimization}
	\subsection{Concept of traffic analyzer: MRTG, PRTG, SNMP, Packet tracer, Wireshark.}

	\pagebreak

\section{Introduction to IPV6}
	\begin{center}(4 Hours/8 Marks)\end{center}
	\subsection{IPv6- Advantages}
	\subsection{Packet formats}
	\subsection{Extension headers}
	\subsection{Transition from IPv4 to IPv6: Dual stack, Tunneling, Header Translation}
	\subsection{Multicasting}

	\pagebreak

\section{Network Security}
	\begin{center}(7 Hours/16 Marks)\end{center}
	\subsection{Properties of secure communication}
	\subsection{Principles of cryptography: Symmetric Key and Public Key}
	\subsection{RSA Algorithm}
	\subsection{Digital Signatures}
	\subsection{Securing e-mail (PGP)}
	\subsection{Securing TCP connections (SSL)}
	\subsection{Network layer security (IPsec, VPN)}
	\subsection{Securing wireless LANs (WEP)}
	\subsection{Firewalls: Application Gateway and Packet Filtering, and IDS}


\end{document}
